---
title: "p8106_stl2137_hw4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lasso2)
library(rpart)
library(rpart.plot)
```

# Question 1

```{r}
### Loading in prostate data
data("Prostate") 
dat_prostate <- Prostate
```

## Part A

```{r}
set.seed(1)

tree_prost_1 <- rpart(formula = lpsa ~., data = dat_prostate)
rpart.plot(tree_prost_1)

### Finding lowest cp 
cp_table <- printcp(tree_prost_1)
#plotcp(tree_prost_1)

minErr <- which.min(cp_table[,4])

# minimum cross-validation error
tree_prost_3 <- prune(tree_prost_1, cp = cp_table[minErr, 1])

# 1SE rule
tree_prost_4 <- prune(tree_prost_1, cp = 
                        cp_table[cp_table[,4] < cp_table[minErr, 4] + cp_table[minErr, 5], 1][1])

rpart.plot(tree_prost_3)
rpart.plot(tree_prost_4)
```

The tree size that corresponds to the lowest cross-validation error is 8. This is not the same tree size as the one obtained using the 1 SE rule, as the tree size is 3. 

## Part B

Pick one of the terminal nodes, and interpret the information displayed.

```{r}
plotcp(tree_prost_1)
```

Based off the leftmost value for which the mean lies below the horizontal line in the cp plot, a cp = 0.1 and a tree size of 3 should be utilized. 

```{r}
set.seed(1)

final_tree_prost <- rpart(lpsa ~ ., data = dat_prostate,
                          control = rpart.control(cp = 0.1))

rpart.plot(final_tree_prost)
```

(c) Perform bagging and report the variable importance.

```{r}

```

